Supplementary material to: Drug efficacy prediction in cell lines

### 2019-01-21 docs/supplement.Rnw
implemented VB and EB steps for conjugate and non-conjugate inv. Gaussian and inv. Gamma model

	In addition to the inverse Gaussian model for the prior variances $\gamma_d^2$ (introduced in the MD):
	$$
  \gamma_d^2 \sim \mathcal{IG}(\theta_d, \lambda_d),
  $$
	we investigate a generalisation of the models in \cite{kpogbezan_empirical_2017} and \cite{leday_gene_2017}, where we draw the $\gamma_d^2$ from an inverse Gamma distribution; :
	$$
	\gamma_d^2 \sim \Gamma^{-1}(\alpha_d/2, \lambda_d/2).
	$$
	For the sake of comparability, we have rescaled the inverse Gamma parameters and reparametrised both distributions as generalized inverse Gaussians:
	\begin{align*}
  \gamma_d^2 & \sim \mathcal{GIG}(-1/2, \lambda_d \theta_d^{-2}, \lambda_d), \\
  \gamma_d^2 & \sim \mathcal{GIG}(-\alpha_d/2, 0, \lambda_d).
  \end{align*}
	The two models coincide if $\theta_d \to \infty$ and $\alpha_d = 1$.

	\subsection{Estimation}
	The variational distributions are as follows:
	\begin{align*}
  q(\bm{\beta}) & \overset{D}{=} \prod_{d=1}^D \mathcal{N}_p (\bm{\mu}_d, \bm{\Sigma}_d), \\
  q(\bm{\gamma}^2) & \overset{D}{=} \prod_{d=1}^D \mathcal{GIG}\(-\frac{p+\alpha_d}{2}, \lambda_d \theta_d^{-2}, \delta_d\), \\
  q(\bm{\sigma}^2) & \overset{D}{=} \prod_{d=1}^D \Gamma^{-1} \(\frac{df + 1}{2}, \zeta_d\),
  \end{align*}
	where $\alpha_d=1$ and $\theta_d \to \infty$ in the inverse Gaussian and inverse Gamma models, respectively. Furthermore, $df=n + p$ in the conjugate model and $df=n$ in the non-conjugate model. The variational parameters contain cyclic dependencies and are, in the conjugate setting, iteratively updated by:
	\begin{align*}
  \bm{\Sigma}_d^{(h+1)} & = (a_d^{(h)})^{-1} (\X \tr \X + b_d^{(h)} \I)^{-1}, \\
  \bm{\mu}_d^{(h+1)} & = (\X \tr \X + b_d^{(h)} \I)^{-1} \X \tr \y_d, \\
  \delta_d^{(h+1)} & = a_d^{(h)} \bmu_d \tr \bmu_d + \lambda_d, \\ 
  \zeta_d^{(h+1)} & = \frac{1}{2} \bigg[\mathbf{y}_d \tr \mathbf{y}_d -2 \mathbf{y}_d \tr \X \bm{\mu}_d^{(h+1)} + \trace ( \X \tr \X \bm{\Sigma}_d^{(h+1)}) + (\bm{\mu}_d^{(h+1)}) \tr \X \tr \X \bm{\mu}_d^{(h+1)} \\
  & \,\,\,\,\,\,\,\,\,\,\,\,\,\,\,\,\,\,\,\,\,\,\,\,\, + b_d^{(h+1)} \trace ( \bm{\Sigma}_d^{(h+1)}) + b_d^{(h+1)} (\bm{\mu}_d^{(h+1)}) \tr \bm{\mu}_d^{(h+1)}\bigg].
  \end{align*}
	while in the non-conjugate setting, we use the following updating equations:
	\begin{align*}
  \bm{\Sigma}_d^{(h+1)} & = (a_d^{(h)})^{-1} (\X \tr \X + b_d^{(h)} (a_d^{(h)})^{-1} \I)^{-1}, \\
  \bm{\mu}_d^{(h+1)} & = (\X \tr \X + b_d^{(h)} (a_d^{(h)})^{-1} \I)^{-1} \X \tr \y_d, \\
  \delta_d^{(h+1)} & = \bmu_d \tr \bmu_d + \lambda_d, \\ 
  \zeta_d^{(h+1)} & = \frac{1}{2} \[\mathbf{y}_d \tr \mathbf{y}_d -2 \mathbf{y}_d \tr \X \bm{\mu}_d^{(h+1)} + \trace ( \X \tr \X \bm{\Sigma}_d^{(h+1)}) + (\bm{\mu}_d^{(h+1)}) \tr \X \tr \X \bm{\mu}_d^{(h+1)} \].
  \end{align*}
  Here, we set
  \begin{align}
  a_d^{(h)} & =\E_{Q^{(h)}}(\sigma_d^{-2})=(df+1)/(2 \zeta_d^{(h)}), \nonumber \\
  b_d^{(h)} & = \E_{Q^{(h)}}(\gamma_d^{-2}) = \sqrt{\frac{\lambda_d}{\theta_d^2 \delta_d^{(h)}}} \frac{K_{\frac{p + \alpha_d -2}{2}} \( \sqrt{\lambda_d \delta_d^{(h)}}/\theta_d \)}{K_{\frac{p+\alpha_d}{2}} \( \sqrt{\lambda_d \delta_d^{(h)}}/\theta_d \)} + \frac{p+\alpha_d^{(h)}}{\delta_d^{(h)}}. \label{eq:auxvar1}
  \end{align}
  Note that in the inverse Gamma model, the computation of (\ref{eq:auxvar1}) simplifies to $b_d^{(h)}=(p+\alpha_d)/\delta_d^{(h)}$.

  \subsection{Empirical Bayes}
  To avoid subjectivity in our choice of prior, we estimate the hyper-parameters with empirical Bayes. The empirical Bayes procedure differs for the two prior models. In the inverse Gaussian case, we parametrise $\theta_d = (\mathbf{c}_d \tr \bm{\alpha})^{-1}$ to allow for the inclusion of drug covariates in the empirical Bayes estimation. The estimating equations are:
  \begin{align*}
  \bm{\alpha} & = \[ \mathbf{C} \tr \diag (\lambda_d e_d^{(l)}) \mathbf{C} \]^{-1} \diag (\lambda_d) \mathbf{1}_{D \times 1}, \\
  \lambda_d & = \[ b_d^{(l)} + e_d^{(l)} (\mathbf{c}_d \tr \bm{\alpha})^2 - 2\mathbf{c}_d \tr \bm{\alpha} \]^{-1},
  \end{align*}
  where $e_d^{(l)} = (b_d^{(l)} + \delta_d^{(l)}/p) \cdot \delta_d^{(l)} (\theta_d^{(h)})^2/\lambda_d^{(l)}$. Solving the equations is done by iteratively reweighted least squares of responses $(e_d^{(l)})^{-1}$ on predictors $\mathbf{c}_d$ with weights $\lambda_d e_d^{(l)}$. If we use one common $\lambda=\lambda_d$ then the solution is in closed form:
  \begin{align*}
	\bm{\alpha}^{(l+1)} & = \[ \mathbf{C} \tr \diag(e_d^{(l)}) \mathbf{C} \]^{-1} \mathbf{C} \tr \mathbf{1}_{D \times 1}, \\
	\lambda^{(l+1)} & = \[ \sum_{d=1}^D b_d^{(l)} - (\bm{\alpha}^{(l+1)}) \tr \mathbf{C} \tr \mathbf{1} \]^{-1} D.
	\end{align*}
	This corresponds to the model in the MD.

  In the inverse Gamma model we use the following empirical Bayes procedure: Suppose we have a partitioning of our drugs into drug classes and we let prior variance components $\gamma_d^2$ and $\gamma^2_{d'}$ be drawn from the same distribution if drug $d$ and drug $d'$ are from the same class. Or simply put: $\gamma^2_{d}, \gamma^2_{d'} \sim \Gamma^{-1}(\alpha_c/2, \lambda_c/2)$ if $d, d' \in \text{class}_c$. Empirical bayes estimation of the $\alpha_c$ and $\lambda_c$ by maximisation of the marginal likelihood amounts to iteratively solving the estimating equations:
  \begin{subequations}
    \begin{align}
      \log \(\frac{\lambda_c}{2}\) - \psi\(\frac{\alpha_c}{2}\) - \frac{\sum_{d \in \text{class}_c} e^{(l)}_d}{|\text{class}_c|} & = 0, \label{eq:estequation1}\\
      \frac{\alpha_c}{\lambda_c} - \frac{\sum_{d \in \text{class}_c} b_d^{(l)}}{|\text{class}_c|} & = 0, \label{eq:estequation2}
    \end{align}
  \end{subequations}
  where $e^{(l)}_d = \log \lambda_d^{(l)} - \psi ( \alpha_d^{(l)}/2) - \log 2$. We solve (\ref{eq:estequation2}) for $\lambda_c$, plug it into (\ref{eq:estequation1}) and bound the result using \cite{alzer_inequalities_1997} to obtain an interval $( a_c^*, 2a_c^* )$ containing the solution to (\ref{eq:estequation1}), where:
  $$
  a_c^* = \left\{ \frac{\sum_{d \in \text{class}_c} e^{(l)}_d}{|\text{class}_c|} + \log \[ \frac{\sum_{d \in \text{class}_c} b_d^{(l)}}{|\text{class}_c|} \]  \right\}^{-1}.
  $$
  With this interval it is straightforward to find $\alpha^{(l+1)}_c$ by any root-finding algorithm and plugging the solution into (\ref{eq:estequation2}) to find $\lambda^{(l+1)}_c$.

@@

	\subsection{Evidence lower bound}
	We monitor convergence through the evidence lower bound (ELBO). To compute the evidence lower bound after updating the EB parameters from $\alpha_d^{(l)}, \theta_d^{(l)}, \lambda_d^{(l)}$ to $\alpha_d^{(l + 1)}, \theta_d^{(l + 1)}, \lambda_d^{(l + 1)}$, we compute $\text{ELBO}^{(l + 1)} = \sum_{d=1}^D \text{ELBO}_d^{(l + 1)}$, with, in the conjugate case:
	\begin{align*}
	\text{ELBO}_d^{(l + 1)} & = \frac{1}{2} \log |\bm{\Sigma}_d^{(l)}| - \frac{a_d^{(l)}}{2} \[ \y_d \tr \y_d - 2\y_d \tr \X \bm{\mu}_d^{(l)} + (\bm{\mu}_d^{(l)}) \tr \X \tr \X \bm{\mu}_d^{(l)} + \trace ( \X \tr \X \bm{\Sigma}_d^{(l)})\] \\
	& - \frac{df + 1}{2} \log \zeta_d^{(l)} + \frac{p + \alpha_d^{(l)}}{4} \log \lambda_d^{(l)} - \frac{p + \alpha_d^{(l)}}{2} \log \theta_d^{(l)} - \frac{p+\alpha_d^{(l)}}{4} \log \delta_d^{(l)} \\
	& + \log K_{\frac{p+\alpha_d^{(l)}}{2}}\(\sqrt{\lambda_d^{(l)} \delta_d^{(l)}}/\theta_d^{(l)}\) + \frac{b_d^{(l)}}{2} \[ \delta_d^{(l)} - a_d^{(l)} (\bm{\mu}_d^{(l)}) \tr \bm{\mu}_d^{(l)} - a_d^{(l)} \trace (\bm{\Sigma}_d^{(l)}) \] \\
	& + \frac{\lambda_d^{(l)} e_d^{(l)}}{2(\theta_d^{(l)})^2} + \frac{(\alpha_d^{(l)} - \alpha_d^{(l + 1)})h_d^{(l)}}{2} - \log K_{\frac{\alpha_d^{(l + 1)}}{2}}\(\lambda_d^{(l + 1)}/\theta_d^{(l + 1)}\) + \frac{\alpha_d^{(l + 1)}}{2} \log \theta_d^{(l)} \\
	& - \frac{\lambda_d^{(l+1)} e_d^{(l)}}{2 (\theta_d^{(l + 1)})^2} - \frac{\lambda_d^{(l + 1)} b_d^{(l)}}{2},
	\end{align*}
	where $h_d^{(l)} = \log \delta_d^{(l)} - \psi [(p + \alpha_d^{(l)})/2] - \log 2$ in the inverse Gamma model and
	$$
	h_d^{(l)} = \frac{1}{2}\log \delta_d^{(l)} - \frac{1}{2}\log \lambda_d^{(l)} + \log \theta_d^{(l)} + \frac{\partial}{\partial \nu} \log K_{\nu} (\sqrt{\lambda_d^{(l)} \delta_d^{(l)}}/\theta_d^{(l)}) \bigg|_{\nu = \frac{p + \alpha_d^{(l)}}{2}}
	$$
	in the inverse Gaussian case. Note that in the inverse Gaussian case, the computationally expensive $h_d^{(l)}$ terms never have to be calculated, since we have $\alpha_d^{(l)} - \alpha_d^{(l + 1)}=0$. The non-conjugate ELBO simply exchanges the $a_d^{(l)} (\bm{\mu}_d^{(l)}) \tr \bm{\mu}_d^{(l)}$ and $a_d^{(l)} \trace (\bm{\Sigma}_d^{(l)})$ terms with $(\bm{\mu}_d^{(l)}) \tr \bm{\mu}_d^{(l)}$ and $\trace (\bm{\Sigma}_d^{(l)})$, respectively. To compute the ELBO after a VB update we simply let $\alpha_d^{(l)}, \theta_d^{(l)}, \lambda_d^{(l)}=\alpha_d^{(l + 1)}, \theta_d^{(l + 1)}, \lambda_d^{(l + 1)}$.

>>>>

	In addition to the inverse Gaussian model for the prior variances $\gamma_d^2$ (introduced in the MD):
	$$
  \gamma_d^2 \sim \mathcal{IG}(\theta_d, \lambda_d),
  $$
	we investigate a generalisation of the models in \cite{kpogbezan_empirical_2017} and \cite{leday_gene_2017}, where we draw the $\gamma_d^2$ from an inverse Gamma distribution; :
	$$
	\gamma_d^2 \sim \Gamma^{-1}(\eta_d/2, \lambda_d/2).
	$$
	For the sake of comparability, we have rescaled the inverse Gamma parameters and reparametrised both distributions as generalized inverse Gaussians:
	\begin{align*}
  \gamma_d^2 & \sim \mathcal{GIG}(-1/2, \lambda_d \theta_d^{-2}, \lambda_d), \\
  \gamma_d^2 & \sim \mathcal{GIG}(-\eta_d/2, 0, \lambda_d).
  \end{align*}
	The two models coincide if $\theta_d \to \infty$ and $\eta_d = 1$.

	\subsection{Estimation}
	The variational distributions are as follows:
	\begin{align*}
  q(\bm{\beta}) & \overset{D}{=} \prod_{d=1}^D \mathcal{N}_p (\bm{\mu}_d, \bm{\Sigma}_d), \\
  q(\bm{\gamma}^2) & \overset{D}{=} \prod_{d=1}^D \mathcal{GIG}\(-\frac{p+\eta_d}{2}, \lambda_d \theta_d^{-2}, \delta_d\), \\
  q(\bm{\sigma}^2) & \overset{D}{=} \prod_{d=1}^D \Gamma^{-1} \(\frac{df + 1}{2}, \zeta_d\),
  \end{align*}
	where $\eta_d=1$ and $\theta_d \to \infty$ in the inverse Gaussian and inverse Gamma models, respectively. Furthermore, $df=n + p$ in the conjugate model and $df=n$ in the non-conjugate model. The variational parameters contain cyclic dependencies and are, in the conjugate setting, iteratively updated by:
  \begin{align*}
  \bm{\Sigma}_d^{(h+1)} & = (a_d^{(h)})^{-1} (\X \tr \X + c_d^{(h)} \I)^{-1}, \\
  \bm{\mu}_d^{(h+1)} & = (\X \tr \X + c_d^{(h)} \I)^{-1} \X \tr \y_d, \\
  \delta_d^{(h+1)} & = b_d^{(h)} (c_d^{(h)})^{-1} \bmu_d \tr \bmu_d + \lambda_d, \\ 
  \zeta_d^{(h+1)} & = \frac{1}{2} \bigg[\mathbf{y}_d \tr \mathbf{y}_d -2 \mathbf{y}_d \tr \X \bm{\mu}_d^{(h+1)} + \trace ( \X \tr \X \bm{\Sigma}_d^{(h+1)}) + (\bm{\mu}_d^{(h+1)}) \tr \X \tr \X \bm{\mu}_d^{(h+1)} \\
  & \,\,\,\,\,\,\,\,\,\,\,\,\,\,\,\,\,\,\,\,\,\,\,\,\, + \frac{b_d^{(h+1)} - c_d^{(h+1)} a_d^{(h+1)}}{1 - a_d^{(h+1)}} \trace ( \bm{\Sigma}_d^{(h+1)}) + \frac{b_d^{(h+1)} - c_d^{(h+1)} a_d^{(h+1)}}{1 - a_d^{(h+1)}} (\bm{\mu}_d^{(h+1)}) \tr \bm{\mu}_d^{(h+1)}\bigg].
  \end{align*}
  Here, we set
  \begin{align}
  a_d^{(h)} & =\E_{Q^{(h)}}(\sigma_d^{-2})=(df+1)/(2 \zeta_d^{(h)}), \nonumber \\
  b_d^{(h)} & = \E_{Q^{(h)}}(\gamma_d^{-2}) = \sqrt{\frac{\lambda_d}{\theta_d^2 \delta_d^{(h)}}} \frac{K_{\frac{p + \eta_d -2}{2}} \( \sqrt{\lambda_d \delta_d^{(h)}}/\theta_d \)}{K_{\frac{p+\eta_d}{2}} \( \sqrt{\lambda_d \delta_d^{(h)}}/\theta_d \)} + \frac{p+\eta_d^{(h)}}{\delta_d^{(h)}}, \label{eq:auxvar1} \\
  c_d^{(h)} & = \begin{cases*}
  b_d^{(h)} & in the conjugate setting \\
  b_d^{(h)}/a_d^{(h)} & in the non-conjugate setting .
  \end{cases*} \nonumber
  \end{align}
  Note that in the inverse Gamma model, the computation of (\ref{eq:auxvar1}) simplifies to $b_d^{(h)}=(p+\eta_d)/\delta_d^{(h)}$.

  \subsection{Empirical Bayes}
  To avoid subjectivity in our choice of prior, we estimate the hyper-parameters with empirical Bayes. The empirical Bayes procedure differs for the two prior models. In the inverse Gaussian case, we parametrise $\theta_d = (\mathbf{c}_d \tr \bm{\alpha})^{-1}$ to allow for the inclusion of drug covariates in the empirical Bayes estimation. The estimating equations are:
  \begin{align*}
  \bm{\alpha} & = \[ \mathbf{C} \tr \diag (\lambda_d e_d^{(l)}) \mathbf{C} \]^{-1} \mathbf{C} \tr \diag (\lambda_d) \mathbf{1}_{D \times 1}, \\
  \lambda_d & = \[ b_d^{(l)} + e_d^{(l)} (\mathbf{c}_d \tr \bm{\alpha})^2 - 2\mathbf{c}_d \tr \bm{\alpha} \]^{-1},
  \end{align*}
  where $e_d^{(l)} = \E_{Q^{(l)}}(\gamma_d^2) = [b_d^{(l)} - (p + \eta_d^{(h)})/\delta_d^{(l)})] \cdot \delta_d^{(l)} (\theta_d^{(h)})^2/\lambda_d^{(l)}$. Solving the equations is done by iteratively reweighted least squares of responses $(e_d^{(l)})^{-1}$ on predictors $\mathbf{c}_d$ with weights $\lambda_d e_d^{(l)}$. If we use one common $\lambda=\lambda_d$ then the solution is in closed form:
  \begin{align*}
	\bm{\alpha}^{(l+1)} & = \[ \mathbf{C} \tr \diag(e_d^{(l)}) \mathbf{C} \]^{-1} \mathbf{C} \tr \mathbf{1}_{D \times 1}, \\
	\lambda^{(l+1)} & = \[ \sum_{d=1}^D b_d^{(l)} - (\bm{\alpha}^{(l+1)}) \tr \mathbf{C} \tr \mathbf{1} \]^{-1} D.
	\end{align*}
	This corresponds to the model in the MD.

  In the inverse Gamma model we use the following empirical Bayes procedure: Suppose we have a partitioning of our drugs into drug classes and we let prior variance components $\gamma_d^2$ and $\gamma^2_{d'}$ be drawn from the same distribution if drug $d$ and drug $d'$ are from the same class. Or simply put: $\gamma^2_{d}, \gamma^2_{d'} \sim \Gamma^{-1}(\eta_c/2, \lambda_c/2)$ if $d, d' \in \text{class}_c$. Empirical bayes estimation of the $\eta_c$ and $\lambda_c$ by maximisation of the (approximated) marginal likelihood amounts to iteratively solving the estimating equations:
  \begin{subequations}
    \begin{align}
      \psi\(\frac{\eta_c}{2}\) & = \log \lambda_c - \frac{\sum_{d \in \text{class}_c} e^{(l)}_d}{|\text{class}_c|} - \log 2, \label{eq:estequation1}\\
      \lambda_c & = \eta_c \cdot |\text{class}_c| \cdot \(\sum_{d \in \text{class}_c} b_d^{(l)}\)^{-1}, \label{eq:estequation2}
    \end{align}
  \end{subequations}
  until convergence, where $e^{(l)}_d = \E_{Q^{(l)}}(\log \gamma_d^2) = \log \delta_d^{(l)} - \psi [ (p + \eta_d^{(l)})/2] - \log 2$. 
  We propose to do that by first solving (\ref{eq:estequation2}) for $\lambda_c$, plugging it into (\ref{eq:estequation1}) and bounding the result using \cite{alzer_inequalities_1997} to obtain an interval $( \eta_c^*, 2\eta_c^* )$ containing the solution to (\ref{eq:estequation1}), where:
  $$
  \eta_c^* = \left\{ \frac{\sum_{d \in \text{class}_c} e^{(l)}_d}{|\text{class}_c|} + \log \[ \frac{\sum_{d \in \text{class}_c} b_d^{(l)}}{|\text{class}_c|} \]  \right\}^{-1}.
  $$
  With this interval it is straightforward to find $\alpha^{(l+1)}_c$ by any root-finding algorithm and plugging the solution into (\ref{eq:estequation2}) to find $\lambda^{(l+1)}_c$. We may use $\eta_c^*$ as a starting value.

@@

	\subsection{Evidence lower bound}
	We monitor convergence through the evidence lower bound (ELBO). To compute the evidence lower bound after updating the EB parameters from $\eta_d^{(l)}, \theta_d^{(l)}, \lambda_d^{(l)}$ to $\eta_d^{(l + 1)}, \theta_d^{(l + 1)}, \lambda_d^{(l + 1)}$, we compute $\text{ELBO}^{(l + 1)} = \sum_{d=1}^D \text{ELBO}_d^{(l + 1)}$, with, in the inverse Gaussian case:
	\begin{align*}
	\text{ELBO}_d^{(l + 1)} & = \frac{1}{2} \log |\bm{\Sigma}_d^{(l)}| - \frac{a_d^{(l)}}{2} \[ \y_d \tr \y_d - 2\y_d \tr \X \bm{\mu}_d^{(l)} + (\bm{\mu}_d^{(l)}) \tr \X \tr \X \bm{\mu}_d^{(l)} + \trace ( \X \tr \X \bm{\Sigma}_d^{(l)})\] \\
	& + \frac{b_d^{(l)}}{2} \[ \delta_d^{(l)} - (a_d^{(l)} + 1 - b_d^{(l)}/c_d^{(l)})(\bm{\mu}_d^{(l)}) \tr \bm{\mu}_d^{(l)} - (a_d^{(l)} + 1 - b_d^{(l)}/c_d^{(l)}) \trace (\bm{\Sigma}_d^{(l)}) \] \\
	& - \frac{df + 1}{2} \log \zeta_d^{(l)} + \frac{p + 1}{4} \log \lambda_d^{(l)} - \frac{p + 1}{2} \log \theta_d^{(l)} - \frac{p+1}{4} \log \delta_d^{(l)} \\
	& + \log K_{\frac{p+1}{2}}\(\sqrt{\lambda_d^{(l)} \delta_d^{(l)}}/\theta_d^{(l)}\) + \frac{\lambda_d^{(l)} e_d^{(l)}}{2(\theta_d^{(l)})^2} \\
	& + \frac{\lambda_d^{(l+1)}}{\theta_d^{(l+1)}} + \frac{1}{2}\log \lambda_d^{(l+1)} - \frac{\lambda_d^{(l+1)} e_d^{(l)}}{2 (\theta_d^{(l + 1)})^2} - \frac{\lambda_d^{(l + 1)} b_d^{(l)}}{2}.
	\end{align*}

	In the inverse Gamma setting, we have:
	\begin{align*}
	\text{ELBO}_d^{(l + 1)} & = \frac{1}{2} \log |\bm{\Sigma}_d^{(l)}| - \frac{a_d^{(l)}}{2} \[ \y_d \tr \y_d - 2\y_d \tr \X \bm{\mu}_d^{(l)} + (\bm{\mu}_d^{(l)}) \tr \X \tr \X \bm{\mu}_d^{(l)} + \trace ( \X \tr \X \bm{\Sigma}_d^{(l)})\] \\
	& + \frac{b_d^{(l)}}{2} \[ \delta_d^{(l)} - (a_d^{(l)} + 1 - b_d^{(l)}/c_d^{(l)})(\bm{\mu}_d^{(l)}) \tr \bm{\mu}_d^{(l)} - (a_d^{(l)} + 1 - b_d^{(l)}/c_d^{(l)}) \trace (\bm{\Sigma}_d^{(l)}) \]  \\
	& - \frac{df + 1}{2} \log \zeta_d^{(l)} - \frac{p + \eta_d^{(l)}}{2} \log \delta_d^{(l)} + \frac{\eta_d^{(l)}}{2} e_d^{(l)} + \frac{\eta_d^{(l)}}{2} \log 2 + \log \Gamma \( \frac{p + \eta_d^{(l)}}{2} \) \\
	& - \log \Gamma \( \frac{\eta_d^{(l + 1)}}{2} \) - \frac{\eta_d^{(l+1)}}{2} e_d^{(l)} - \frac{\eta_d^{(l+1)}}{2} \log 2 + \frac{\eta_d^{(l + 1)}}{2} \log \lambda_d^{(l+1)} - \frac{\lambda_d^{(l + 1)} b_d^{(l)}}{2}.
	\end{align*}
  Note that the computation of $e_d^{(l)}$ differs between the inverse Gaussian and Gamma models. Furthermore, to compute the ELBO after a VB update we simply let $\eta_d^{(l)}, \theta_d^{(l)}, \lambda_d^{(l)}=\eta_d^{(l + 1)}, \theta_d^{(l + 1)}, \lambda_d^{(l + 1)}$.
